---
title: "Final project: Regression"
author: "Jose Antonio Alvarez Ocete"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For fractions:
```{r}
library(MASS)
```

## R Markdown

```{r}
#import data
data <- read.table("data/battery.dat")

p <- 5

#set up X
X <- data.matrix(data[,1:p])
n <- length(X[,1])
Z <- cbind(rep(1,n), X)
Y <- data.matrix(data[,p+1])

names <- c('','Z1','Z2','Z3','Z4','Z5')
colnames(Z)<- names

sampleMean<-function(X, n) {
  Ones <- rep(1,n)
  return (1/n * t(X)%*%Ones)
}

sampleCovariance<-function(X, n, sample_mean) {
  Ones <- rep(1,n)
  return (1/(n-1) * t(X - Ones%*%t(sample_mean))%*%(X - Ones%*%t(sample_mean)))
}
```

(1) Find the least square estimate beta_hat

```{r}
# least square estimates
beta_hat <- solve(t(Z)%*%Z)%*%t(Z)%*%Y
rownames(beta_hat) <- c('beta0','beta1','beta2','beta3','beta4','beta5')
beta_hat
```

(2) Find the R^2 statistic

```{r}
# R^2 statistic
R_square <- 1 - sum((Y - Z%*%beta_hat)^2)/sum((Y-mean(Y))^2)
R_square
```

(3) Find sigma_hat_square and estimated Cov(beta_square)

```{r}
# sigma_hat_square
sigma_hat_square <- sum((Y - Z%*%beta_hat)^2)/(n-p-1)
cat('Sigma hat square:', sigma_hat_square, fill=TRUE)

# estimated covariance of hat{beta}
covariance_hat <- sigma_hat_square * solve(t(Z)%*%Z)
cat('Covariance Hat:', fill=TRUE)
covariance_hat
```

NOT USED:

```{r}
# t-test for single coefficient
# H_0: beta_j = 0, H_a: beta_j != 0

j <- 1
t_stat <- (beta_hat[j+1] - 0)/sqrt(sigma_hat_square * solve(t(Z)%*%Z)[j+1,j+1])
t_stat

alpha <- 0.05
cval_t <- qt(1-alpha/2, n-p-1)
cval_t
```

(4) 95% confidence interval for the beta:

```{r}
# One-at-a-time confidence interval for beta_j

for (j in 0:p) {
  cat('Beta hat', j,': [',
    beta_hat[j+1] - qt(1-alpha/2, n-p-1)*sqrt(sigma_hat_square * solve(t(Z)%*%Z)[j+1,j+1]),
    ',',
    beta_hat[j+1] + qt(1-alpha/2, n-p-1)*sqrt(sigma_hat_square * solve(t(Z)%*%Z)[j+1,j+1]),
    ']', fill=TRUE)
}
```

(5) 95% simultaneous confidence intervals for all betas based on the confidence region

```{r}
# confidence region based simultaneous confidence intervals 

for (j in 0:p) {
cat('Beta hat', j,': [',
      beta_hat[j+1] - sqrt((p+1)*qf(1-alpha, p+1, n-p-1))*sqrt(sigma_hat_square * solve(t(Z)%*%Z)[j+1,j+1]),
      ',',
      beta_hat[j+1] + sqrt((p+1)*qf(1-alpha, p+1, n-p-1))*sqrt(sigma_hat_square * solve(t(Z)%*%Z)[j+1,j+1]),
      ']', fill=TRUE)
}
```

(6) 95% simultaneous confidence intervals for all betas based on the Bonferroni correction

```{r}
# Bonferroni correction based simultaneous confidence intervals

for (j in 0:p) {
  cat('Beta hat', j,': [',
    beta_hat[j+1] - qt(1-alpha/(2*(p+1)), n-p-1)*sqrt(sigma_hat_square * solve(t(Z)%*%Z)[j+1,j+1]),
    ',',
    beta_hat[j+1] + qt(1-alpha/(2*(p+1)), n-p-1)*sqrt(sigma_hat_square * solve(t(Z)%*%Z)[j+1,j+1]),
    ']', fill=TRUE)
}
```

(7) Test H_0: beta_1 = beta_2 = 0 at significance level alpha = 0.05

```{r}
# F-test
# H_0: beta_1 = beta_2 = 0

C <- cbind(rep(0,p), diag(p))

df_1 <- qr(C)$rank # df_1: rank of matrix R

f_stat <- t(C%*%beta_hat)%*%solve(C%*%solve(t(Z)%*%Z)%*%t(C))%*%(C%*%beta_hat)
f_stat

cval_f <- sigma_hat_square*df_1*qf(1-alpha, 2, n-p-1)
cval_f

# (equivalent) F-test by comparing residuals

# fit the reduced model
beta_hat_reduced <- solve(t(Z[,1])%*%Z[,1])%*%t(Z[,1])%*%Y
beta_hat_reduced

f_stat_reduced <- ((sum((Y - Z[,1]%*%beta_hat_reduced)^2) - sum((Y - Z%*%beta_hat)^2))/2)/sigma_hat_square
f_stat_reduced
```

(8) 95% confidence interval for the mean response given z_0

```{r}
# confidence interval for z_0^T beta

X_mean <- sampleMean(X,n)
z_0 <- rbind(1, X_mean)

cat('C.I. for the mean response z_0[',
  t(z_0)%*%beta_hat - sqrt(sigma_hat_square)*sqrt(t(z_0)%*%solve(t(Z)%*%Z)%*%z_0)*qt(1-alpha/2, n-p-1),
  ',',
  t(z_0)%*%beta_hat + sqrt(sigma_hat_square)*sqrt(t(z_0)%*%solve(t(Z)%*%Z)%*%z_0)*qt(1-alpha/2, n-p-1),
  ']')
```

(9) 95% confidence interval for a new response given z_0

```{r}
# prediction interval for Y_0 = z_0^T beta + epsilon_0

cat('Prediction interval for Y_0 = z_0^T beta + epsilon_0: [',
  t(z_0)%*%beta_hat - sqrt(sigma_hat_square)*sqrt(1+t(z_0)%*%solve(t(Z)%*%Z)%*%z_0)*qt(1-alpha/2, n-p-1),
  ',',
  t(z_0)%*%beta_hat + sqrt(sigma_hat_square)*sqrt(1+t(z_0)%*%solve(t(Z)%*%Z)%*%z_0)*qt(1-alpha/2, n-p-1),
  ']')
```

NOT USED:

```{r}
# Confidence Region for (beta_1, beta_2)^T

center <- beta_hat[2:3]
es<-eigen(R%*%solve(t(Z)%*%Z)%*%t(R))
e1<-es$vec %*% diag(sqrt(es$val))
r1<-sqrt(df_1*cval_f*sigma_hat_square)
theta<-seq(0,2*pi,len=250)
v1<-cbind(r1*cos(theta), r1*sin(theta))
pts<-t(center - (e1%*%t(v1)))
plot(pts,type="l",main="Confidence Region for (beta_1, beta_2)^T",xlab="beta_1",ylab="beta_2",asp=1,
     xlim = c(-0.3,0.1),ylim=c(-0.1,0.04))
segments(0,center[2],center[1],center[2],lty=2) # highlight the center
segments(center[1],0,center[1],center[2],lty=2)
arrows(-0.3,0,0.1,0)
arrows(0,-0.1,0,0.05)

th2<-c(0,pi/2,pi,3*pi/2,2*pi)   #adding the axis
v2<-cbind(r1*cos(th2), r1*sin(th2))
pts2<-t(center-(e1%*%t(v2)))
segments(pts2[3,1],pts2[3,2],pts2[1,1],pts2[1,2],lty=3)  
segments(pts2[2,1],pts2[2,2],pts2[4,1],pts2[4,2],lty=3)
```